---
title: 'R Programming: Take Home MidTerm'
author: "Denis Vrdoljak"
date: "Dec 4, 2019"

output: html_notebook
---

QUESTION 1

Complete the following questions, then submit this R Notebook with your work on Camino.

Load the data in the file "fillna.txt", replace the NA's with "unknown" and save to a new file called "cleaned.txt"
```{r}
## Student code goes here
#set a wd

data = read.table("fillna.txt")
data[is.na(data)] <- "unknown"
head(data)
write.table(data, file = "cleaned.txt", row.names = FALSE, col.names = FALSE)

## Student code goes here
```

QUESTION 2

Load the table fillna.txt and 
a) create a vector of NA counts by column
b) create a vector of NA counts by row

```{r}
## Student code goes here

mydata = read.table("fillna.txt")
count_col_na <- colSums(is.na(mydata))
print("Vector of NA counts by column: ")
count_col_na
count_row_na <- rowSums(is.na(mydata))
print("Vector of NA counts by row: ")
count_row_na

## Student code goes here
```

QUESTION 3

Load the mtcars dataset, and show the structure of the table.

```{r}
## Student code goes here
data("mtcars")
str(mtcars)

## Student code goes here
```


QUESTION 4
Load the airquality dataset and show the first 7 rows. (PART A)

```{r}
## Student code goes here
data("airquality")
head(airquality,7)


## Student code goes here
```


... and show the last 4 rows of data (PART B)

```{r}
## Student code goes here
tail(airquality,4)


## Student code goes here
```


Show data for the month of May (PART C)

```{r}
## Student code goes here
library("dplyr")
mayData <- airquality %>%
  filter(Month == 5)

mayData

## Student code goes here
```

QUESTION 5

Load the Orange dataset, and perform the following analyses. Organize your work and present your answers with separate R Notebook cells.

a. Do a simple scatterplot of Age vs circumference.
b. Make a stripchart of age
c. Change the graph from part (a) as follows:
1. Use red dots
2. Label the graph from exercise one as:
-- Title = Tree Age vs circumference
-- X axis = age
-- Y axis = circumference of tree

** STUDENT CODE CELLS SHOULD GO HERE. Create your own cells! **

Load the Orange dataset, and perform the following analyses. Organize your work and present your answers with separate R Notebook cells.
```{r}
#downloading dataset
data("Orange")
#showing data
head(Orange)
```

a. Do a simple scatterplot of Age vs circumference.
```{r} 

plot(Orange$age,Orange$circumference)
 
```

b. Make a stripchart of age
```{r}
stripchart(Orange$age)

```

c. Change the graph from part (a) as follows:
1. Use red dots
2. Label the graph from exercise one as:
-- Title = Tree Age vs circumference
-- X axis = age
-- Y axis = circumference of tree
```{r}
 
plot(Orange$age, Orange$circumference,
     main = "Tree Age vs circumference",
     xlab = "Age", 
     ylab = "circumference of Tree",
     col  = "orange")
```

QUESTION 6
Using the Orange Dataset again, create a graph that shows the distribution of values of the circumference column

```{r}
## Student code goes here

x <- Orange$circumference
h<-hist(x, breaks=25, xlab="Circumference",
   main="Distribution of the Circumference")
xfit<-seq(min(x),max(x),length=40)
yfit<-dnorm(xfit,mean=mean(x),sd=sd(x))
yfit <- yfit*diff(h$mids[1:2])*length(x)
lines(xfit, yfit, col="blue", lwd=2)
lines(density(Orange$circumference), col="blue", lwd=2)

## Student code goes here
```