---
title: 'R Programming: Take Home MidTerm'
author: "Denis Vrdoljak"
date: "Dec 4, 2019"

output: html_notebook
---

Complete the following questions, then submit this R Notebook with your work on Camino.

QUESTION 1

Load the dataset arquality and show days im June when temperature was higher then 70..
```{r}
## Student code goes here
data(airquality)
sub <- subset(airquality, Month == 6 & Temp > 70)
sub
## Student code goes here
```

QUESTION 2

Load the table fillna.txt and 
a) create a vector of NA counts by column
b) create a vector of NA counts by row

```{r}
## Student code goes here
mydata = read.table("fillna.txt")
count_col_na <- colSums(is.na(mydata))
print("Vector of NA counts by column: ")
count_col_na
count_row_na <- rowSums(is.na(mydata))
print("Vector of NA counts by row: ")
count_row_na

## Student code goes here
```

QUESTION 3

Load the orange dataset, and show the structure of the table.

```{r}
## Student code goes here
data("Orange")
str(Orange)
## Student code goes here
```

QUESTION 4
Load the ggplot2 library. Next, load the 'diamonds' dataset. (Note you may use dplyr for these questions, but do not need to.)

Plot the diamonds' clarity by price (PART A)

```{r}
## Student code goes here

library(ggplot2)

ggplot(diamonds, aes(factor(clarity), price, fill=clarity)) + geom_boxplot() + ggtitle("Diamond Clarity by Price") + xlab("Clarity") + ylab("Diamond Price ($)") + coord_cartesian(ylim=c(0,6500))
## Student code goes here
```


Show the last 4 rows of data (PART B)

```{r}
## Student code goes here
print(tail(diamonds, n=4))
## Student code goes here
```


Show the median price  (PART C)

```{r}
## Student code goes here
print(median(diamonds$price))
## Student code goes here
```

QUESTION 5

Load the Orange dataset, and perform the following analyses. Organize your work and present your answers with separate R Notebook cells.

a. Do a simple scatterplot of Age vs circumference.

```{r}
## Student code goes here
data("Orange")
plot(Orange$age,Orange$circumference)
## Student code goes here
```
b. Make a stripchart of age

```{r}
## Student code goes here
stripchart(Orange$age,
           main = "Stripchart of Age",
           ylab = "Age")
## Student code goes here
```
c. Change the graph from part (a) as follows:
1. Use orange dots
2. Label the graph exercise one as:
-- Title = Tree Age vs circumference
-- X axis = Age
-- Y axis = circumference of Tree

```{r}
## Student code goes here
plot(Orange$age, Orange$circumference,
     main = "Tree Age vs Circumference",
     xlab = "Age", 
     ylab = "Circumference",
     col  = "orange")
## Student code goes here
```

QUESTION 6
Using the Diamond dataset again, create a graph that shows the distribution of values for the diamond prices column. 
```{r}

qplot(x = diamonds$price,data = diamonds,
        binwidth = 100,
      breaks = c(0,500,1000,1500,2000,3000,4000,5000,7500,10000,15000,20000),
        xlab = "Diamond Price",
        ylab = "Count",
        main = "Diamond Price Distribution",
        col = I("green"))   
 
```

