---
title: "R Notebook"
output: html_notebook
---

Required library, we are going to use dataset that had student's GRE score and their admittance

```{r}
library(naivebayes)
library(dplyr)
library(psych)
```

Load the data we need ----> 

```{r}
data <- read.csv(file.choose(), header = T)
str(data)
```
convert variables to factor variable
```{r}
data$rank <- as.factor(data$rank)
data$admit <- as.factor(data$admit)
```
use random seeding and create training 80% and testing 20% data sets
```{r}
set.seed(1234)
ind <- sample(2, nrow(data), replace = T, prob = c(0.8, 0.2))
train <- data[ind == 1,]
test <- data[ind == 2,]
```

build the NB model, as admit depends on all variables here, use training data

```{r}
model <- naive_bayes(admit ~ ., data = train)
model
```
Test for accuracy in training data set, 70% accurate
```{r}
p1 <- predict(model, test)
(tab1 <- table(p1, test$admit))
sum(diag(tab1)) / sum(tab1)
```
Decision tree with party, seems like GPA is the most determining variable....
```{r}
library(party)
tree <- ctree(admit ~., data = train)
tree
plot(tree)
```
Test for accuracy, 72%
```{r}
pt <- predict(tree, train)
(tab2 <- table(pt, train$admit))
sum(diag(tab2)) / sum(tab2)
```





