---
title: 'R Programming: Take Home MidTerm'
author: "Denis Vrdoljak"
date: "Dec 4, 2019"
---


PART 1


1.1. Load the titanic.csv data

```{r}
data <- read.csv("titanic.csv")
head(data)
```



1.2. Get rid of the following columns: 'Name', 'Ticket', 'PassengerId', 'Cabin'

```{r}
data$Name <- NULL
data$Ticket <- NULL
data$PassengerId <- NULL
data$Cabin <- NULL

head(data)
```



1.3. Replace the ‘Sex column with a binary column for whether the passenger is ‘Male’

```{r}
#Male is 1 and female is 0
levels(data$Sex) <- c(0,1)

head(data)

```



1.4. Show how many NA’s are in each column

```{r}
na_counts <- colSums(is.na(data))
na_counts

```



1.5. Which Column has the most NA’s? How many does it have?

```{r}

print("Column with the most NA's")
max_na<- subset(na_counts,na_counts==(max(na_counts)))
max_na

```



1.6. Create a bar chart to show NA counts by column

```{r}

Na_plot<- barplot(na_counts, main="NA Count by column",col="royalblue",ylim =c(0,200))

```




1.7. Show/plot the distribution of ages for each gender. Is the average (mean) age of Males higher? Is the median higher?

```{r}

#Clean out NA ages first and store as data_clean for this part
data_clean <- na.omit(data)

colSums(is.na(data_clean))

male <- subset(data_clean, Sex == 1)
female <- subset(data_clean, Sex == 0)

#Male Average Age
print (paste("Male Average Age:",  mean(male$Age)))

#Female Average Age
print (paste("Female Average Age:", mean(female$Age)))

#Plot distribution of ages for each gender as boxplot
plot(data_clean$Sex, data_clean$Age)


```
1 represents males, while 0 represent females. 

The mean age of male is 30.72664. The mean age of female is 27.91571. So the average age of males is greater. 

The median age of males is higher as shown by the box plot.




1.8. Which port of origin (embarked column) had the highest average fare? Show/plot your results.

```{r}

library(ggplot2)
data1 <- data[rowSums(is.na(data)) == 0,]
head(data1)
ggplot(data1, aes(x = factor(Embarked), y = Fare)) +
  theme_bw() +
  geom_boxplot() +
  stat_summary(fun.y=mean, geom="point", shape=21, size=4) +
  geom_hline(aes(yintercept=35), colour='red', linetype='dashed', lwd=0.5) +
  labs(title = "Embarked vs Fare", subtitle = "C = Cherbourg; Q = Queenstown; S = Southampton", x = "Port of Origin", y = "Fare")



#Using original data again
#Gather total unique port of origins
ports <- unique(data$Embarked)
ports

#Average fare for C
C <- subset(data, Embarked == "C")
C_fare <- mean(C$Fare)
C_fare


#Average fare for 
Q <- subset(data, Embarked == "Q")
Q_fare <- mean(Q$Fare)
Q_fare

#Average fare for C
S <- subset(data, Embarked == "S")
S_fare <- mean(S$Fare)
S_fare

#Max fare
max_fare <- max(c(C_fare,Q_fare,S_fare))
max_fare

max_port = subset(data, sum(C$Fare) == max_fare)
max_port

print(paste("The higest average fare is ", max(c(C_fare,Q_fare,S_fare))))

```


The port of origin with the higest average fare is "C" at 59.95414.



1.9. Is there a statistically significant difference in average fares between the genders?

```{r}
#Using cleaned data
#Males average fare
print (paste("Males average fare", mean(male$Fare)))

#Females average fare
print (paste("Females average fare", mean(female$Fare)))

#t-test to 
t.test(male$Fare, female$Fare)
```

Although the average fares between the genders seem to vary, the t-test results in a p-value of 4.362e-06, which is less than 0.05.
As such, we fail to reject the null hypothesis that there is no difference between the average fares of the two genders.


PART 2

2.1. Load the bitly data from data.gov as a data frame.

```{r}
#Load JSON as a dataframe
library(jsonlite)
library(data.table)
library(stringr)
library(tidyr)
data2 <- stream_in(file("bitly_usagov_example.txt"))
head(data2)
```

2.2. How many records are there?

```{r}

print(paste("Number of records: ", nrow(data2)))
```


2.3. In the City column (‘cy’), how many are NA?

```{r}
na_counts <- sum(is.na(data2$cy))
print(paste("Number of NAs in the City column: ", na_counts))
```


2.4. How many countries are present in this data? Which country (‘c’ column) has the most records? How many does it have?

```{r}
#Number of countries present
print(paste("There are ", length(unique(data2$c)), " unique countries"))

#Country with most records

print(paste("The ", names(which(table(data2$c) == max(table(data2$c)))), "  has the most number of records at ", max(table(data2$c))))

```


2.5. How many records are from Russia?

```{r}
countries <- table(data2$c)

print(paste(countries[names(countries)=="RU"], " records are from Russia.")) 
```


2.6. How many records do NOT list USA

```{r}

print(paste( "Including NA records, there are ",
length(data2$c)-countries[names(countries)=="US"],
" records that are not US."))
```


2.7. How many records have 'America' in the timezone, but Country not in US?

```{r}
print(paste("There are ",
sum(data2$tz %like% "America" & data2$c != "US", na.rm = TRUE),
" such records"))
```


2.8. How many records point to cia.gov?

```{r}
print(paste("There are ",
sum(data2$r %like% "cia.gov" | data2$hh %like% "cia.gov" | data2$u %like% "cia.gov", na.rm = TRUE), 
" such records"))
```


2.9. Fill empty records with “UNKNOWN” and NA’s with “MISSING”. Create a new csv. (Submit this csv with your R Notebook.)

```{r}
data3 <- data2
data3[is.na(data3)] <- "MISSING"
data3[data3 == ""] <- "UNKNOWN"

#Fix ll column to character
data3$ll <- as.character(data3$ll)

write.csv(data3, file = "BetterData.csv")

```


2.10. Show the top 10 timezones (‘tz’ column) in a bar chart, with a legend, and properly labeled x,y axes and title.

```{r}
top_10 <- sort(table(data2$tz),decreasing=TRUE)[1:10]
print(paste("Top10 timezones", " "))
top_10

barplot(top_10, main = "Bitly",legend.text = "Top 10 Timezones", xlab = "Time Zones", ylab = "Frequency")

```

